---
title: "intro-to-LTMFNpackage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro-to-LTMFNpackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(LTMFNpackage)
```
#Introduction to the LTMFNpackage

Hello and welcome to LTMFNpackage! This package was designed to regroup useful functions for the Long-Term Monitoring of Fish Tracking Networks project (LTMFTN). It mainly includes functions useful for:

1. Data entry quality control
2. Data processing (loading, filtering, cleaning, and formatting)
3. Some helper functions for using YAPS

First, let's load the package:
```{r}
library(LTMFNpackage)
```

##Data entry quality control

This package contains a suite of data quality control functions to be used on manual data entry. These functions are specifically designed for use with our internal excel database. This database is formatted with multiple sheets (LIST OF SHEETS) to enter different data types

The following functions are used for quality control
check_dataentry: general function to check data entry
append_to_database: add quality controlled data to the clean database
import_database_xl: function to import excel database

There are also functions available for checking individual sheets. These are not designed as part of the regular data pipeline, but are made available for performing individual checks
check_equipment_log
check_fish
check_fykes
check_angling
check_cast_netting
check_range_test
check_gps_records.

To perform quality control according to the data management protocol:
1. Place all data entry excel files into the raw data folder on the lab Surface Pro ([EXACT DIRECTORY])
2. call the function check_dataentry(folder_path = "[EXACT DIRECTORY]")
    This function will check that the surface pro is being used, and that the proper working directory is set [EXACT DIRECTORY]. For standardized data management, ensure this is the case. For the purposes of this example, the correct folder_path and directory will not be used
```{r}
setwd("~/Documents/Pedersen2023/LTMFTN/LTMFNpackage/data-raw/data_qc_example")
check_dataentry(folder_path = "raw/")

```

##Data Processing 


##Yaps helper functions
This package contains helper functions to facilitate analysis using the YAPS package.

Let's say we imported a raw download from the VUE software 
```{r}
data(vue_det)
head(vue_det)
```

First, let's format the receiver detections data to make it compatible with the yaps package.
This formats the detections data to have the columns

* ts: time stamp in POSIXct
tag: tag as the last 4 digits of the tag ID number in integer
epo: the time of the detecton in unix time (integer)
frac: the time fraction of the detection (sub-second) (double)
serial: the serial number of the receiver (integer)
```{r}
yaps_ex_data <- format_receiverlog_yaps(receiverLog = vue_det, 
                                        full = TRUE, 
                                        starttime = "2023-08-15 09:45:00", 
                                        endtime = "2023-08-15 10:00:00")
head(yaps_ex_data)

```
